\section{Discussion}
\label{sec:discussion}

\subsection{Key Findings}

\textbf{Prompt design rivals model scale for faithfulness.} Switching Concise$\rightarrow$Detailed raises faithfulness by up to 12pp for capable models---comparable to upgrading Qwen3-32B to Qwen3.5-397B. For Qwen3.5-397B the template switch alone lifts mean faithfulness from 0.84 to 0.93. As prompt changes require no hardware upgrades and add no inference overhead, they should be the first optimization lever.

\textbf{MMR improves coverage but requires context management.} MMR improves retrieval recall and nDCG but can dilute faithfulness for smaller models: GPT-OSS-20B drops from 0.82 to 0.79 faithfulness despite nDCG@5 rising from 0.70 to 0.79, indicating the 20B model cannot reliably separate relevant evidence from peripheral passages. Practitioners should pair MMR with a larger generation model or a reranking step.

\textbf{Chunk size effects are secondary.} 500 vs.\ 1000 character chunks differ by 2--5pp on most metrics, vs.\ 9--12pp for retrieval strategy and up to 12pp for template choice. Retrieval and prompt optimization should precede chunk-parameter tuning.

\subsection{Metric Divergence}

Lexical metrics (METEOR, ROUGE-L, BLEU) diverge from semantic metrics (BERTScore, faithfulness) across templates: Detailed lowers METEOR by 10--22pp while raising faithfulness by up to 12pp, because SQuAD references are short noun phrases while Detailed answers include supporting context. BERTScore F1 varies by only 1--3pp, confirming semantic stability. Metric selection should follow the application's verbosity tolerance.

\subsection{Threats to Validity}

\textbf{Internal.} LLM-based metrics depend on the Qwen3.5-397B critic, which may favour same-family outputs; we report IR and NLG metrics as complementary signals.

\textbf{External.} Results are obtained on SQuAD, a single-domain extractive QA dataset; abstractive or domain-specific tasks may differ. The framework supports additional datasets (RagPerf-Wikipedia-NQ, T2-RagBench, FinQA).

\textbf{Statistical.} Bootstrap 95\% CIs (Table~\ref{tab:stat-significance}) are tight ($\pm 0.01$); paired Wilcoxon tests confirm headline template and retrieval effects at $p<0.05$ for five of seven models. Sub-0.01 differences between top-tier models fall within CI overlap.

\textbf{Reproducibility.} We use default server temperatures without fixed seeds; the reproducibility bundle captures model/dependency versions and git hash, but exact numerical replication may vary slightly due to non-deterministic GPU operations.

\subsection{Limitations}

The study excludes reranking, HyDE query expansion, and embedding-model comparisons. All quality assessments are automated; no human evaluation is included.
